Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation

### 2022-08-23 Paper1.tex
Final version paper 23.08.2022

\begin{equation}
\label{eq:FloquetGF} 
\underline{G}_{mn}(\omega) =\int \dd t_{\text{rel}} \int_{-\tau/2}^{\tau/2} \frac{dt_{\text{av}}}{\tau} \ee^{\ii[\left(\omega+m\Omega\right) t -\left( \omega+n\Omega\right)t^{\prime}]} \underline{G}(t,t'),
\end{equation}

known as the \emph{Keldysh-Floquet} GF. The integration variables $t_{\text{rel}} = t-t^{\prime}$ and $t_{\text{av}} = (t+t^{\prime})/2$ are the relative and average times. Within the rest of this work, we denote a Floquet-represented matrix either as $X_{mn}$ with explicit indices, or use a boldface letter $\bf{X}$. The underline indicates the overall \emph{Keldysh} structure

\begin{equation}\label{eq:Keld-structure}
\underline{{\bf G}} \equiv 
\begin{pmatrix}
{\bf G}^{\text{R}} & {\bf G}^{\text{K}}\\
{\bf 0}         & {\bf G}^{\text{A}} \\
\end{pmatrix},
\end{equation}

which contains the \emph{retarded}, \emph{advanced} and \emph{Keldysh} components $\textbf{G}^{\text{R,A,K}}$, where $\textbf{G}^{\text{A}}=(\textbf{G}^{\text{R}})^{\dagger}$. The \emph{Keldysh} component is defined as $\textbf{G}^{\text{K}} \equiv \textbf{G}^{>} + \textbf{G}^{<}$, with $\textbf{G}^{\lessgtr}$ as \emph{lesser} and \emph{greater} components~\cite{schw.61,keld.65,ra.sm.86,ha.ja}.

@@

\begin{equation}\label{eq:non-int_InvGF} 
\underline{G}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) =  \underline{g}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) - \sum_{\rho\in\{\text{l},\text{r}\}}v^2_{\rho}\underline{g}_{\text{b},\rho}(\omega_n;\epsilon)\delta_{mn}
\end{equation}

with $v_{\text{l}/\text{r}}$ as lead-layer hybridization, $\omega_n \equiv \omega + n\Omega$, $n\in\mathbbm{Z}$, and

\begin{equation}\label{eq:inv_non-int_lat_GF_comps}
\begin{split}
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{R}}_{mn} & = \left( \omega_n + i0^{+} -\varepsilon_{\text{c}} \right)\delta_{mn} - \varepsilon_{mn}(\epsilon,\overline{\epsilon}), \\
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{K}}_{mn} & = 0
\end{split}
\end{equation}

as the noninteracting Green's function of the isolated layer, whose inverse Keldysh component is negligible in the steady state. The Floquet dispersion relation $\varepsilon_{mn}$ for the periodic field in a hyper-cubic lattice is~\cite{ts.ok.08}

\begin{equation}
\label{eq:Floquet_disp}
\varepsilon_{mn}(\epsilon,\overline{\epsilon}) = \begin{cases}
\epsilon J_{m-n}(\aamp) & m-n:\text{even}, \\
\ii\overline{\epsilon}J_{m-n}(\aamp) & m-n:\text{odd},
\end{cases}
\end{equation} 
% 
where $J_n$ denotes the $n$-th order Bessel function of the first kind, with the argument $\aamp$ defined in Sec.~\ref{sec:Model}. The  GF $\underline{g}_{\text{b},\text{l}/\text{r}}$ on the $d$-dimensional boundary of the decoupled leads,  consisting of a single layer coupled to a semi-infinite reservoir in the wide-band limit with broadening $\Gamma_{\text{l}/\text{r}}$ (cf. Fig.~\ref{fig:setup}) is given by

\begin{align}
g^{\text{R}}_{\text{b},\text{l}/\text{r}}(\omega;\epsilon) &= \frac{1}{\omega-\varepsilon_{\text{l}/\text{r}}(\epsilon)+\ii\Gamma_{\text{l}/\text{r}}},\label{eq:retarded_bath_GF} \\
g^{\text{K}}_{\text{b},\text{l}/\text{r}}(\omega;\epsilon) &= 2\ii \iim[g_{\text{b},\text{l}/\text{r}}^{\text{R}}(\omega;\epsilon)][1-2f(\omega,\mu_{\text{l}/\text{r}},\beta)], \label{eq:keldysh_bath_GF}
\end{align}

where $\varepsilon_{\text{l}/\text{r}}(\epsilon)=\varepsilon_{\text{l}/\text{r}} +\frac{t_{\text{l}/\text{r}}}{t^\ast}\epsilon$ is the dispersion, $f(\omega,\mu_{\text{l}/\text{r}},\beta)=1/[\ee^{\beta(\omega-\mu_{\text{l}/\text{r}})}+1]$ the Fermi-Dirac distribution function at inverse temperature $\beta \equiv 1/T$, $\varepsilon_{\text{l}/\text{r}}$ the onsite energy and $t_{\text{l}/\text{r}}$ the hopping within the $d$-dimensional layer of the lead.

@@

The Keldysh components of the non-interacting phonon GF $\kel{D}_{\text{ph}}(t,t^{\prime})$ are given by~\cite{ao.ts.14}
\begin{align}\label{eq:Ph_Prop_time}
\begin{split} 
D^{\text{R}}_{\text{ph}}(t,t^{\prime}) & = -\ii \theta(t-t^{\prime}) \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega), \\
D^{>}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \left[ 1 + b(\omega) \right] \\
D^{<}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \ b(\omega), \\
\end{split}, 
\end{align}
where $b(\omega)=1/(\ee^{\beta\omega}-1)$ is the Bose-Einstein distribution function at inverse temperature $\beta$. We focus on acoustic phonons, with spectral function $A_{\text{ph}}(\omega) = (\omega/\omega^{2}_{\text{ph}}) \ee^{-| \omega|/\omega_{\text{ph}}}$,  $\omega_{\text{ph}}$ being a soft cutoff frequency~\cite{pi.li.21}. The retarded and Keldysh components of the e-ph SE are easily extracted from  Eq.~\eqref{eq:backbone_e-ph_SE} and can be found in Ref.~\cite{ma.ga.22u}.

@@

We briefly describe now the self-consistent F-DMFT scheme used.
(i) We start from an initial guess for the electron SE $\kel{\mat{\Sigma}}(\omega)$ and set $\kel{\mat \Sigma}_{\text{e-ph}}(\omega)=0$. (ii) Then we compute the local electron GF as

\begin{equation}\label{eq:Lat_LocGF}
\begin{split}
\kel{\mat G}_{\text{loc}}(\omega) &= \int \dd\epsilon \int \dd\overline{\epsilon} \ \rho(\epsilon,\overline{\epsilon}) \times \\ 
&\times \left[\kel{\mat G}^{-1}_{0}(\omega,\epsilon,\overline{\epsilon}) - \kel{\mat\Sigma}(\omega) - \kel{\mat \Sigma}_{\text{e-ph}}(\omega) \right]^{-1}.
\end{split}
\end{equation}

(iii) Using Eq.~\eqref{eq:backbone_e-ph_SE}, we obtain the phonon contribution to the SE $\kel{\mat{\Sigma}}_{\text{e-ph}}(\omega)$.
(iv) The problem is mapped onto a single impurity plus bath, whose hybridization function
is given by

\begin{equation}\label{eq:imp_Dyson_eq}
\kel{\mat{\Delta}}(\omega) = \kel{\mat{g}}^{-1}_{0,\text{site}}(\omega) - \kel{\mat{G}}^{-1}_{\text{loc}}(\omega) - \kel{\mat{\Sigma}}(\omega),
\end{equation}

where $\kel{\mat{g}}^{-1}_{0,\text{site}}(\omega)$ is defined as in Eq.~\eqref{eq:inv_non-int_lat_GF_comps} with $\varepsilon_{mn}(\epsilon,\overline{\epsilon})=0$. (v) The nonequilibrium many-body impurity problem is solved according to the procedure in Sec.~\ref{sec:amea}, leading to the new
 $\kel{\mat{\Sigma}}(\omega)$. (vi) We insert the electron and e-ph SEs into step (ii) and iterate the steps (ii)-(vi) until convergence.

@@

Due to the hybridization with the leads, the local DOS of the Hubbard layer only features a pseudogap
 $\Delta_{\text{pg}} \approx 4$.
All simulations are carried out with the dimensionless factor $\alpha\equiv t_{r}E_0/\Omega^{2}<0.5$ for which the FDSA is justified~\cite{so.do.18}\footnote{In Ref.~\cite{so.do.18} the factor is defined as $\alpha=t_{c}E_0/\Omega^{2}$ and $t_{\text{c}}=1$. Setting $t_{r}=1$, we define for consistency $\alpha$ as in the main text.} and first-order photon absorption processes are dominant. Whenever the electron-phonon interaction is switched on, we take $g=0.8$ and unless stated otherwise $\omega_{\text{ph}}=0.1$.

@@

\begin{figure}[b] 
% \captionsetup{width=\textwidth} 
\includegraphics[width=\linewidth]{./figures_Paper1/j_vs_omega_mu1_sweep_v_E0_2_eph.pdf}
\caption{(Color online) (a) Time-averaged steady-state current $j$ and (b) double occupancy $N_{\text{D}}$ as a function of $\Omega$, for $v=0.5,0.6,0.7$, without and with e-ph interaction. Default parameters are specified at the beginning of Sec.~\ref{sec:results}.} 
\label{fig:j_vs_omega_mu1_sweep_v_E0_2_eph}
\end{figure}

@@

Experimentally, the electric field amplitude considered in this paper is several orders of magnitude bigger than the sunlight one, however it is possible to achieve such amplitudes via THz pulses ~\cite{mu.we.18} ~\footnote{We do not consider smaller electric field amplitudes $E_0$ since DMFT only converges slowly for such a choice of parameters. Moreover, the observables' magnitude becomes comparable to our numerical precision.}. Moreover, in order to address the occurrence of II in realistic photovoltaic setups, one should extend such simplified model to multiorbital systems \cite{pe.be.19} in a multilayer setup, to model oxide heterostructures~\cite{as.bl.13,pe.be.19} where the correlated region consists of multiple layers and photoexcited carriers are separated via an electric field gradient. Another step forward would be to take into account the effect of impurity scattering and the feedback of the electrons onto the phonon dynamics with a self-consistent treatment.

>>>>

\begin{equation}
\label{eq:FloquetGF} 
\underline{G}_{mn}(\omega) =\int \dd t_{\text{rel}} \int_{-\tau/2}^{\tau/2} \frac{dt_{\text{av}}}{\tau} \ee^{\ii[\left(\omega+m\Omega\right) t -\left( \omega+n\Omega\right)t^{\prime}]} \underline{G}(t,t'),
\end{equation}
known as the \emph{Keldysh-Floquet} GF. The integration variables $t_{\text{rel}} = t-t^{\prime}$ and $t_{\text{av}} = (t+t^{\prime})/2$ are the relative and average times. Within the rest of this work, we denote a Floquet-represented matrix either as $X_{mn}$ with explicit indices, or use a boldface letter $\bf{X}$. The underline indicates the overall \emph{Keldysh} structure
%
\begin{equation}\label{eq:Keld-structure}
\underline{{\bf G}} \equiv 
\begin{pmatrix}
{\bf G}^{\text{R}} & {\bf G}^{\text{K}}\\
{\bf 0}         & {\bf G}^{\text{A}} \\
\end{pmatrix},
\end{equation}
%
which contains the \emph{retarded}, \emph{advanced} and \emph{Keldysh} components $\textbf{G}^{\text{R,A,K}}$, where $\textbf{G}^{\text{A}}=(\textbf{G}^{\text{R}})^{\dagger}$. The \emph{Keldysh} component is defined as $\textbf{G}^{\text{K}} \equiv \textbf{G}^{>} + \textbf{G}^{<}$, with $\textbf{G}^{\lessgtr}$ as \emph{lesser} and \emph{greater} components~\cite{schw.61,keld.65,ra.sm.86,ha.ja}.

@@

\begin{equation}\label{eq:non-int_InvGF} 
\underline{G}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) =  \underline{g}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) - \sum_{\rho\in\{\text{l},\text{r}\}}v^2_{\rho}\underline{g}_{\text{b},\rho}(\omega_n;\epsilon)\delta_{mn}
\end{equation}
with $v_{\text{l}/\text{r}}$ as lead-layer hybridization, $\omega_n \equiv \omega + n\Omega$, $n\in\mathbbm{Z}$, and 
%
\begin{equation}\label{eq:inv_non-int_lat_GF_comps}
\begin{split}
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{R}}_{mn} & = \left( \omega_n + i0^{+} -\varepsilon_{\text{c}} \right)\delta_{mn} - \varepsilon_{mn}(\epsilon,\overline{\epsilon}), \\
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{K}}_{mn} & = 0
\end{split}
\end{equation}
%
as the noninteracting Green's function of the isolated layer, whose inverse Keldysh component is negligible in the steady state. The Floquet dispersion relation $\varepsilon_{mn}$ for the periodic field in a hyper-cubic lattice is~\cite{ts.ok.08} 
%
\begin{equation}
\label{eq:Floquet_disp}
\varepsilon_{mn}(\epsilon,\overline{\epsilon}) = \begin{cases}
\epsilon J_{m-n}(\aamp) & m-n:\text{even}, \\
\ii\overline{\epsilon}J_{m-n}(\aamp) & m-n:\text{odd},
\end{cases}
\end{equation} 
% 
where $J_n$ denotes the $n$-th order Bessel function of the first kind, with the argument $\aamp$ defined in Sec.~\ref{sec:Model}. The  GF $\underline{g}_{\text{b},\text{l}/\text{r}}$ on the $d$-dimensional boundary of the decoupled leads,  consisting of a single layer coupled to a semi-infinite reservoir in the wide-band limit with broadening $\Gamma_{\text{l}/\text{r}}$ (cf. Fig.~\ref{fig:setup}) is given by
%  
\begin{align}
g^{\text{R}}_{\text{b},\text{l}/\text{r}}(\omega;\epsilon) &= \frac{1}{\omega-\varepsilon_{\text{l}/\text{r}}(\epsilon)+\ii\Gamma_{\text{l}/\text{r}}},\label{eq:retarded_bath_GF} \\
g^{\text{K}}_{\text{b},\text{l}/\text{r}}(\omega;\epsilon) &= 2\ii \iim[g_{\text{b},\text{l}/\text{r}}^{\text{R}}(\omega;\epsilon)][1-2f(\omega,\mu_{\text{l}/\text{r}},\beta)], \label{eq:keldysh_bath_GF}
\end{align} 
%
where $\varepsilon_{\text{l}/\text{r}}(\epsilon)=\varepsilon_{\text{l}/\text{r}} +\frac{t_{\text{l}/\text{r}}}{t^\ast}\epsilon$ is the dispersion, $f(\omega,\mu_{\text{l}/\text{r}},\beta)=1/[\ee^{\beta(\omega-\mu_{\text{l}/\text{r}})}+1]$ the Fermi-Dirac distribution function at inverse temperature $\beta \equiv 1/T$, $\varepsilon_{\text{l}/\text{r}}$ the onsite energy and $t_{\text{l}/\text{r}}$ the hopping within the $d$-dimensional layer of the lead.

@@

The Keldysh components of the non-interacting phonon GF $\kel{D}_{\text{ph}}(t,t^{\prime})$ are given by~\cite{ao.ts.14}
%
\begin{align}\label{eq:Ph_Prop_time}
\begin{split} 
D^{\text{R}}_{\text{ph}}(t,t^{\prime}) & = -\ii \theta(t-t^{\prime}) \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega), \\
D^{>}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \left[ 1 + b(\omega) \right] \\
D^{<}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \ b(\omega), \\
\end{split}, 
\end{align}
%
where $b(\omega)=1/(\ee^{\beta\omega}-1)$ is the Bose-Einstein distribution function at inverse temperature $\beta$. We focus on acoustic phonons, with spectral function $A_{\text{ph}}(\omega) = (\omega/\omega^{2}_{\text{ph}}) \ee^{-| \omega|/\omega_{\text{ph}}}$,  $\omega_{\text{ph}}$ being a soft cutoff frequency~\cite{pi.li.21}. The retarded and Keldysh components of the e-ph SE are easily extracted from  Eq.~\eqref{eq:backbone_e-ph_SE} and can be found in Ref.~\cite{ma.ga.22u}.

@@

We briefly describe now the self-consistent F-DMFT scheme used.
(i) We start from an initial guess for the electron SE $\kel{\mat{\Sigma}}(\omega)$ and set $\kel{\mat \Sigma}_{\text{e-ph}}(\omega)=0$. (ii) Then we compute the local electron GF as 
%  
\begin{equation}\label{eq:Lat_LocGF}
\begin{split}
\kel{\mat G}_{\text{loc}}(\omega) &= \int \dd\epsilon \int \dd\overline{\epsilon} \ \rho(\epsilon,\overline{\epsilon}) \times \\ 
&\times \left[\kel{\mat G}^{-1}_{0}(\omega,\epsilon,\overline{\epsilon}) - \kel{\mat\Sigma}(\omega) - \kel{\mat \Sigma}_{\text{e-ph}}(\omega) \right]^{-1}.
\end{split}
\end{equation}
% 
(iii) Using Eq.~\eqref{eq:backbone_e-ph_SE}, we obtain the phonon contribution to the SE $\kel{\mat{\Sigma}}_{\text{e-ph}}(\omega)$.
(iv) The problem is mapped onto a single impurity plus bath, whose hybridization function
is given by
%
\begin{equation}\label{eq:imp_Dyson_eq}
\kel{\mat{\Delta}}(\omega) = \kel{\mat{g}}^{-1}_{0,\text{site}}(\omega) - \kel{\mat{G}}^{-1}_{\text{loc}}(\omega) - \kel{\mat{\Sigma}}(\omega),
\end{equation}   
%  
where $\kel{\mat{g}}^{-1}_{0,\text{site}}(\omega)$ is defined as in Eq.~\eqref{eq:inv_non-int_lat_GF_comps} with $\varepsilon_{mn}(\epsilon,\overline{\epsilon})=0$. (v) The nonequilibrium many-body impurity problem is solved according to the procedure in Sec.~\ref{sec:amea}, leading to the new
 $\kel{\mat{\Sigma}}(\omega)$. (vi) We insert the electron and e-ph SEs into step (ii) and iterate the steps (ii)-(vi) until convergence.

@@

Due to the hybridization with the leads, the local DOS of the Hubbard layer only features a pseudogap
 $\Delta_{\text{pg}} \approx 4$.
All simulations are carried out with the dimensionless factor $\alpha\equiv t_{r}E_0/\Omega^{2}<0.5$ for which the FDSA is justified~\footnote{In Ref.~\cite{so.do.18} the factor is defined as $\alpha=t_{c}E_0/\Omega^{2}$ and $t_{\text{c}}=1$. Setting $t_{r}=1$, we define for consistency $\alpha$ as in the main text.} and first-order photon absorption processes are dominant. Whenever the electron-phonon interaction is switched on, we take $g=0.8$ and unless stated otherwise $\omega_{\text{ph}}=0.1$.

@@

\begin{figure}[t] 
% \captionsetup{width=\textwidth} 
\includegraphics[width=\linewidth]{./figures_Paper1/j_vs_omega_mu1_sweep_v_E0_2_eph.pdf}
\caption{(Color online) (a) Time-averaged steady-state current $j$ and (b) double occupancy $N_{\text{D}}$ as a function of $\Omega$, for $v=0.5,0.6,0.7$, without and with e-ph interaction. Default parameters are specified at the beginning of Sec.~\ref{sec:results}.} 
\label{fig:j_vs_omega_mu1_sweep_v_E0_2_eph}
\end{figure}

@@

Experimentally, the electric field amplitude considered in this paper~\footnote{We do not consider smaller electric field amplitudes $E_0$ since DMFT only converges slowly for such a choice of parameters. Moreover, the observables' magnitude becomes comparable to our numerical precision.} is several orders of magnitude larger than that of the sunlight, however it is possible to achieve such amplitudes via THz pulses ~\cite{mu.we.18}. Moreover, in order to address the occurrence of II in realistic photovoltaic setups, one should extend such simplified model to multiorbital systems \cite{pe.be.19} in a multilayer setup, to model oxide heterostructures~\cite{as.bl.13,pe.be.19} where the correlated region consists of multiple layers and photoexcited carriers are separated via an electric field gradient. Another step forward would be to take into account the effect of impurity scattering and the feedback of the electrons onto the phonon dynamics with a self-consistent treatment.
